SIRV-RL: Reinforcement Learning for Optimized Policy Control during Epidemiological Outbreaks in Emerging Market and Developing Economies

### 2024-02-14 paper_draft/main.tex
paper edits, graphs changed

In this paper we optimize the government policies regarding stringency. Therefore, alongside epidemiological data, we use the measures of globally comparable government responses~\cite{Hale2021}. We use the simple SIR model without vital dynamics~\cite{Hethcote1989, Hethcote2008, ALLEN2017128} as it is assumed that the timescale is small enough that it can be neglected~\cite{Cooper2020}. By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we can effectively address the real-world conditions and propose a solution that is both effective and extendable. The current model (SIRV with lockdown) accounts for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{SIRLockdown,NBERw26981,NBERw26867}. Traditional SIR model neglects the time-varying property of parameters like $\beta$ and $\gamma$, but in this paper since, lockdown is a parameter of the proposed model, and it changes with time therefore, we propose a time-dependent SIR model\cite{TimeDependentSIR}. However, the study has limitations; First, the deterministic SIR model (predecessor to our proposed model) fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}. Second, the underreporting of cases during the period selected by our study. Lastly, the reinforcement learning agent should be resistant to how the vaccination rate changes -- or to when it is first introduced because then it scopes the environment for succumbing to wishful thinking which can be  potentially dangerous.

After modelling the disease with lockdown (via stringency index) and vaccination, we try to understand the effects of lockdown on the GDP. Therefore, decisions made by the government regarding the level of lockdown to be enforced plays a role on both the public health outcomes and economic stability during a pandemic. On one hand, stringent lockdown measures can effectively slow the spread of the disease, thereby improving public health outcomes. However, these measures often come at the cost of significant economic disruption, leading to job losses, business closures, and reduced economic growth. On the other hand, relaxing lockdown measures may help to mitigate the economic impact of the pandemic, but could result in increased disease transmission and worsened public health outcomes. In order to capture competing costs within the environment and achieve a balance between health and economic outcomes, we intend to employ reinforcement learning~\cite{Nguyen2022,Bastani2021,Song2022,Ohi2020,PADMANABHAN2021102676}. Not only does the formulation of the model deal better with competing costs, but it also offers more transparency behind the reasoning of the decisions being made in such circumstances. When we conceptualize our problem as a reinforcement learning task, an agent is tasked with making decisions in an environment with the aim of optimizing cumulative rewards (i.e., the total amount of reward it receives over the long run). Simply put, given discrete time steps $t = 0, 1, 2, 3, \dots$, at each time step the agent receives a representation of the environment's state, $s_t \in \altmathcal{S}$, and selects an action $a_t \in \altmathcal{A}(s_t)$, where $\altmathcal{A}(s_t)$ is the set of actions available in state $s_t$, and one step later receives a reward $r_{t+1} \in \altmathcal{R}$ and the state is updated~\cite{sutton2018reinforcement}. The way we define the way these rewards that are given to the agent, makes this decision process more transparent, however, it has its limitations. A universal optimal policy may not suit diverse socioeconomic contexts due to variations in healthcare resources and economic vulnerabilities across countries, regions, or cities and a comprehensive consideration of decision factors, extending beyond pure reinforcement learning results is needed~\cite{Song2022,Dunn2017,Demir2006}.

@@

In summary, by using reinforcement learning augmented with deep learning techniques for the SIRV with lockdown environment, we can better understand the effects of lockdown measures on both public health outcomes and economic stability during a pandemic. However, it is crucial to consider the limitations of this approach and take into account a comprehensive set of decision factors in order to make informed policy decisions that are tailored to specific socioeconomic contexts.

@@

At the peak of an epidemic $R_0$ declines as there are no more susceptible people left in the pool, therefore, $R_e$ (effective reproductive number) comes into play. $R_e$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic, therefore, it takes into account the existing immunity of the system. 
\begin{equation}
  R_e = R_0 \frac{S}{N}
  \label{eq:R_effective}
\end{equation}
$S$ is the number of susceptible people, $N$ is the total population. At the start of an epidemic when everyone is susceptible, $R_e = R_0$ as, $S = N$ (i.e., the whole population is susceptible). $\beta$ and $\gamma$ are also used to define probability of and infectious individual infecting another individual $\beta / (\beta + \gamma)$ and the probability of recovery, $\gamma / (\beta + \gamma)$.

@@

To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define two cost functions \cref{eq:cost_SIR,eq:cost_I} to calibrate the model with the use of huber loss~\cite{huberloss}. We generally calibrate the model using \cref{eq:cost_SIR}. However, sometimes we must strike a balance between modeling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}).
\begin{equation}
  L_{\delta}(y, f(x)) = 
  \begin{cases}
    \frac{1}{2}{(y - f(x))^2} & \text{for } |y - f(x)| \leq \delta, \\
    \delta \cdot (|y - f(x)| - \frac{1}{2}\delta) & \text{otherwise.} \\
  \end{cases}
  \label{eq:huberloss}
\end{equation}
In the above equation, $y$ is the actual data and $f(x)$ is the prediction.
\begin{equation}
  \begin{split}
  \textrm{loss\_SIR} = \textrm{cost\_function\_SIR}(S, \hat{S}, I, \hat{I}, R, \hat{R}) \\
  = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
  \end{split}
  \label{eq:cost_SIR}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = \textrm{cost\_function\_I}(I, \hat{I}) = L_{\delta = 1}(I, \hat{I})
  \label{eq:cost_I}
\end{equation}
Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered. Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown} , we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data. The following parameters and loss is obtained:
\begin{equation}
  \beta_{optimal} = 0.042
  \label{eq:beta_optimal_without_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.024
  \label{eq:gamma_optimal_without_lockdown}
\end{equation}
\begin{equation}
  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.762
  \label{eq:r0_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 85051490.533
  \label{eq:cost_SIR_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 45187665.281
  \label{eq:cost_I_without_lockdown}
\end{equation}
See \cref{fig:SIR_model_IND} to see how the model compares with the actual data.

@@

Lastly, an additional flow from the susceptible population to the recovered population can be shown by adding a vaccination rate $\nu$ in the model. 

@@

  \textrm{loss\_SIR} = 10840360.995

@@

\subsection{Optimizing Window Length for Time-varying Vaccination Rate}\label{optimizing_window_length}
However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time. This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}. Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} i.e., the window length that result in the least loss from \cref{eq:cost_SIR}. Using different window lengths to estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$, we observe the following. It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
\begin{equation}
  \frac{dS}{dt} = -\beta_{optimal}  (1 - s(t)/100)  \frac{S I}{N} - \nu S
  \label{eq:S_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta_{optimal}  (1 - s(t)/100) \frac{S I}{N} - \gamma_{optimal} I
  \label{eq:I_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma_{optimal} I + \nu S
  \label{eq:R_with_lockdown_and_nu_calc_nu}
\end{equation}
The main difference here is instead of supplying these equations with the entire time series, only data lying within the sub-interval of $[{start}, {start} + {window\_length}]$ is given, then in the next step data of the next sub-inteval is provided to optimize for, i.e., $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$. This model gives shows how $\nu$ varies with time. Since, the window length acts like a hyperparameter, we can try different window lengths in order to evaluate which would give us the best results. Therefore, the below graphs show us the loss using \cref{eq:cost_SIR,eq:cost_I}.

@@

From the following graph, although the window length of $10$ is best for all three groups of the population but this window length leads to a bad approximation for the infected population therefore, we choose a window length of $15$.

@@

This shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India \cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}. Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIRV Model with time varying $\nu$.

\subsection{SIRV Model with Lockdown (Time-varying Vaccination Rate)}
Lastly, we present our model using the following set of equations:
\begin{figure}[htbp!]
\begin{tikzpicture}[node distance=1cm, auto,
  >=Latex, 
  every node/.append style={align=center},
  int/.style={draw, minimum size=1cm}]
 \node [int] (S)             {$S$};
 \node [int, right=of S] (I) {$I$};
 \node [int, right=of I] (R) {$R$};
 \coordinate[right=of I] (out);
 \path[->] (S) edge node {$\lambda$} (I)
           (I) edge node {$\gamma$} (out)
           (S) edge[bend left=45] node[above] {$\nu(t)$} (R); % Arrow from S to R with text \nu(t)
\end{tikzpicture}
\end{figure}
\begin{equation}
  \frac{dS}{dt} = -\beta  (1 - s(t)/100)  \frac{S I}{N} - \nu(t) S
  \label{eq:S_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta  (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I + \nu(t) S
  \label{eq:R_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \beta_{optimal} = 0.463
  \label{eq:beta_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.114
  \label{eq:gamma_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \begin{split}
    \overbar{R_0} = 1.546 \\
    \widetilde{R_0} = 1.483 \\
    \text{Mode}(R_0) = 0.734 \\
    \sigma_{R_0} = 0.718 \\
    R_0 \in [0.150, 2.785]
  \end{split}
  \label{eq:r0_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 29116762.926
  \label{eq:cost_SIR_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 658537.443
  \label{eq:cost_I_with_lockdown_and_time_varying_nu}
\end{equation}
See \Crefrange{fig:SIR_model_IND}{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \cref{fig:comparing_costs_IND} to see how the different models compare against each other.

Now, that a relation between $\beta$ and $s(t)$ is set up, it must be investigated how stringency index affects the normalized Gross domestic product (GDP)~\cite{OECDNormalizedGDP,OECDNormalizedGDP2}. To do this a polynomial equation of the third degree is fitted to the data points $(x, y)$ where $x$ is the $s(t)$ stringency at time $t$, and $y$ the normalized GDP and minimize the squared error. For India after fitting a 3 degree polynomial, the following equation is obtained:
\begin{equation}
    \begin{split}
      GDP = -5.96640236 \times 10^{-5} s^{3} + 6.65064332 \times 10^{-3} s^{2} - 2.23109924 \times 10^{-1} s^{1} \\
    + 1.01357226 \times 10^{2}
    \end{split}
    \label{eq:gdp_modelled_with_stringency_IND}
\end{equation}

\subsection{Reinforcement Learning}
Given that the government is an agent that takes decisions in a deterministic environment defined above, we use reinforcement learning to model the competing costs of the environment. This environment is formally as a Markov decision process, and can be described as follows:

>>>>

In this paper we optimize the government policies regarding stringency as it controls both the spread of the disease and the economy. To model the epidemiological data~\cite{WorldometerCorona} we use the simple SIR model without vital dynamics~\cite{Hethcote1989, Hethcote2008, ALLEN2017128} as it is assumed that the timescale is small enough that it can be neglected~\cite{Cooper2020}. By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we effectively model the disease progression. Our model (SIRV with lockdown and time-varying vaccination rate) in addition to the SIR model, accounts for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{Hale2021,SIRLockdown,NBERw26981,NBERw26867}. Eventhough the traditional SIR model is a valuable tool for understanding the spread of infectious diseases, it assumes that parameters like the transmission rate ($\beta$) and recovery rate ($\gamma$) are constant over time, which may not always be the case. In this paper, we propose a more sophisticated approach by introducing a time-dependent SIR model~\cite{TimeDependentSIR}, enabling us to account for the changing dynamics of the pandemic due to factors such as lockdowns and vaccination rates. This proposed model effectively addresses the real-world conditions acts as a solution that is both effective and extendable. However, the study has limitations; First, the deterministic SIR model (predecessor to our proposed model) fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}. Second, the underreporting of cases during the period selected by our study. Lastly, the reinforcement learning agent should be resistant to how the vaccination rate changes -- or to when it is first introduced because then it scopes the environment for succumbing to wishful thinking which can be  potentially dangerous.

After modelling the disease with lockdown (via stringency index) and vaccination, we try to understand the effects of lockdown on the GDP~\cite{StringencyEconomicDecline,Cilloni2020,Arinaminpathy2010}. Therefore, decisions made by the government regarding the level of lockdown to be enforced plays a role on both the public health outcomes and economic stability during a pandemic. On one hand, stringent lockdown measures can effectively slow the spread of the disease, thereby improving public health outcomes. However, these measures often come at the cost of significant economic disruption, leading to job losses, business closures, and reduced economic growth. On the other hand, relaxing lockdown measures may help to mitigate the economic impact of the pandemic, but could result in increased disease transmission and worsened public health outcomes. In order to capture competing costs within the environment and achieve a balance between health and economic outcomes, we intend to employ reinforcement learning~\cite{Nguyen2022,Bastani2021,Song2022,Ohi2020,PADMANABHAN2021102676}. Not only does the formulation of the model deal better with competing costs, but it also offers more transparency behind the reasoning of the decisions being made in such circumstances. When we conceptualize our problem as a reinforcement learning task, an agent is tasked with making decisions in an environment with the aim of optimizing cumulative rewards (i.e., the total amount of reward it receives over the long run). Simply put, given discrete time steps $t = 0, 1, 2, 3, \dots$, at each time step the agent receives a representation of the environment's state, $s_t \in \altmathcal{S}$, and selects an action $a_t \in \altmathcal{A}(s_t)$, where $\altmathcal{A}(s_t)$ is the set of actions available in state $s_t$, and one step later receives a reward $r_{t+1} \in \altmathcal{R}$ and the state is updated~\cite{sutton2018reinforcement}. The way we define the way these rewards that are given to the agent, makes this decision process more transparent, however, it has its limitations. A universal optimal policy may not suit diverse socioeconomic contexts due to variations in healthcare resources and economic vulnerabilities across countries, regions, or cities and a comprehensive consideration of decision factors, extending beyond pure reinforcement learning results is needed~\cite{Song2022,Dunn2017,Demir2006}.

@@

In summary, by using reinforcement learning augmented with deep learning techniques for the SIRV with lockdown and time-varying vaccination rate environment, we can better understand the effects of lockdown measures on both public health outcomes and economic stability during a pandemic. However, it is crucial to consider the limitations of this approach and take into account a comprehensive set of decision factors in order to make informed policy decisions that are tailored to specific socioeconomic contexts.

@@

At the peak of an epidemic there is a decline as there are no more susceptible people left in the pool, therefore, $R_e$ (effective reproductive number) comes into play. $R_e$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic, therefore, it takes into account the existing immunity of the system~\cite{MASSAD2017232}. 
\begin{equation}
  R_e = R_0 \frac{S(t)}{N}
  \label{eq:R_effective}
\end{equation}
$S$ is the number of susceptible people, $N$ is the total population. At the start of an epidemic when everyone is susceptible, $R_e = R_0$ as, $S = N$ (i.e., the whole population is susceptible). $\beta$ and $\gamma$ are also used to define probability of and infectious individual infecting another individual $\beta / (\beta + \gamma)$ and the probability of recovery, $\gamma / (\beta + \gamma)$.

@@

To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define two cost functions \cref{eq:cost_SIR,eq:cost_I} to calibrate the model with the use of huber loss~\cite{huberloss}. We generally calibrate the model using \cref{eq:cost_SIR}. However, sometimes we must strike a balance between modeling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}, where a hyperparameter is selected keeping both these losses in mind).
\begin{equation}
  L_{\delta}(y, f(x)) = 
  \begin{cases}
    \frac{1}{2}{(y - f(x))^2} & \text{for } |y - f(x)| \leq \delta, \\
    \delta \cdot (|y - f(x)| - \frac{1}{2}\delta) & \text{otherwise.} \\
  \end{cases}
  \label{eq:huberloss}
\end{equation}
In the above equation, $y$ is the actual data and $f(x)$ is the prediction.
\begin{equation}
  \begin{split}
  \textrm{loss\_SIR} = \textrm{cost\_function\_SIR}(S, \hat{S}, I, \hat{I}, R, \hat{R}) \\
  = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
  \end{split}
  \label{eq:cost_SIR}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = \textrm{cost\_function\_I}(I, \hat{I}) = L_{\delta = 1}(I, \hat{I})
  \label{eq:cost_I}
\end{equation}
Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered. Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown} , we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data. The following parameters and loss is obtained:
\begin{equation}
  \beta_{optimal} = 0.042
  \label{eq:beta_optimal_without_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.024
  \label{eq:gamma_optimal_without_lockdown}
\end{equation}
\begin{equation}
  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.762
  \label{eq:r0_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 85051490.533
  \label{eq:cost_SIR_without_lockdown}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 45187665.281
  \label{eq:cost_I_without_lockdown}
\end{equation}
See \cref{fig:SIR_model_IND} to see how the model compares with the actual data.

@@

Lastly, an additional flow from the susceptible to recovered population can be shown by adding a vaccination rate $\nu$ in the model. 

@@

  \textrm{loss\_I} = 10840360.995

@@

\subsection{Optimizing Window Length for Time-varying Vaccination Rate}\label{optimizing_window_length}
However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time. This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}. Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} i.e., a window length that results in the least loss from \cref{eq:cost_SIR,eq:cost_I}. Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ then for the next time-step $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$ we again estimate $\nu$ -- so on and so forth. It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
\begin{equation}
  \frac{dS}{dt} = -\beta_{optimal}  (1 - s(t)/100)  \frac{S I}{N} - \nu S
  \label{eq:S_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta_{optimal}  (1 - s(t)/100) \frac{S I}{N} - \gamma_{optimal} I
  \label{eq:I_with_lockdown_and_nu_calc_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma_{optimal} I + \nu S
  \label{eq:R_with_lockdown_and_nu_calc_nu}
\end{equation}
This model gives shows how $\nu$ varies with time. Since, the window length acts like a hyperparameter, we can try different window lengths in order to evaluate which would give us the best results. Therefore, the below graphs show us the loss using \cref{eq:cost_SIR,eq:cost_I}.

@@

The following graph indicates that a window length of $10$ days yields the least overall loss for all three population groups. However, this window length results in a poor approximation for the infected group, which is crucial for accurately modeling the spread of the disease. Consequently, we have decided to use a window length of $15$ days, which provides a more accurate approximation for the infected population while still maintaining reasonable loss for the other groups.

@@

This shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}. Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIRV Model with time varying $\nu$.

\subsection{SIRV Model with Lockdown (Time-varying Vaccination Rate)}
Finally, we integrate the time-varying vaccination rate ($\nu$) into the SIR model that includes lockdown measures, resulting in the following set of equations, which represents our final model:
\begin{figure}[htbp!]
\begin{tikzpicture}[node distance=1cm, auto,
  >=Latex, 
  every node/.append style={align=center},
  int/.style={draw, minimum size=1cm}]
 \node [int] (S)             {$S$};
 \node [int, right=of S] (I) {$I$};
 \node [int, right=of I] (R) {$R$};
 \coordinate[right=of I] (out);
 \path[->] (S) edge node {$\lambda$} (I)
           (I) edge node {$\gamma$} (out)
           (S) edge[bend left=45] node[above] {$\nu(t)$} (R); % Arrow from S to R with text \nu(t)
\end{tikzpicture}
\end{figure}
\begin{equation}
  \frac{dS}{dt} = -\beta  (1 - s(t)/100)  \frac{S I}{N} - \nu(t) S
  \label{eq:S_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta  (1 - s(t)/100) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I + \nu(t) S
  \label{eq:R_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \beta_{optimal} = 0.463
  \label{eq:beta_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.114
  \label{eq:gamma_optimal_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \begin{split}
    \overbar{R_0} = 1.546 \\
    \widetilde{R_0} = 1.483 \\
    \text{Mode}(R_0) = 0.734 \\
    \sigma_{R_0} = 0.718 \\
    R_0 \in [0.150, 2.785]
  \end{split}
  \label{eq:r0_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_SIR} = 29116762.926
  \label{eq:cost_SIR_with_lockdown_and_time_varying_nu}
\end{equation}
\begin{equation}
  \textrm{loss\_I} = 658537.443
  \label{eq:cost_I_with_lockdown_and_time_varying_nu}
\end{equation}
See \Crefrange{fig:SIR_model_IND}{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \Crefrange{fig:comparing_costs_SIR_IND}{fig:comparing_costs_I_IND} to see how the different models compare against each other.

\subsection{Modelling Normalized GDP with Stringency}
Now, that a relation between $\beta$ and $s(t)$ is set up, it must be investigated how stringency index affects the normalized gross domestic product (GDP)~\cite{OECDNormalizedGDP,OECDNormalizedGDP2}. To do this a polynomial equation of the third degree is fitted to the data points $f(x) = a x^3 + b x^2 + c x + d = y$, here, $x$ is the $s(t)$ stringency at time $t$, and $y$ the normalized GDP, and we minimize the squared error to find the values of coefficients $a, b, c, d$. For India after fitting a 3 degree polynomial, the following equation is obtained:
\begin{equation}
    \begin{split}
      \textrm{normalized\_GDP} = -5.96640236 \times 10^{-5} s^{3} + 6.65064332 \times 10^{-3} s^{2} - 2.23109924 \times 10^{-1} s^{1} \\
    + 1.01357226 \times 10^{2}
    \end{split}
    \label{eq:gdp_modelled_with_stringency_IND}
\end{equation}

\subsection{Reinforcement Learning}
Given that the government is an agent that takes decisions in a deterministic environment defined above, we use reinforcement learning to model the competing costs of the environment. This environment is known as a Markov decision process (MDP), and can be described as follows:
% https://arxiv.org/pdf/1708.05866.pdf
\begin{itemize}
    % \item \url{https://jmlr.org/papers/volume22/20-1364/20-1364.pdf} RL Paper that references \url{https://araffin.github.io/slides/rl-tuto-jnrr19/#/13} RL ppt (for our RL notes).
    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP ($(GDP_{predicted} - GDP_{min}) / (GDP_{max} - GDP_{min})$), $R_e$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible ($S$), infected ($I$) and recovered ($R$). The starting states are simply these values at the starting date and no previous actions.
    \item Actions $\altmathcal{A}$: The stringency index variable was analyzed with a sample size of 915. The mean value was approximately $61.965049$, with a standard deviation of $17.669831$. The minimum value was $31.480000$, while the maximum value reached $96.300000$. And the differences between two consecutive stringencies had a mean of $-0.070919$, and standard deviation of $1.427145$, with the minimum being $-14.360000$ and maximum $16.670000$. Based on this we define the discrete action space. There are 7 actions for the agent, it can keep the stringency index same, reduce/increase by 2.5, reduce/increase by 5, and reduce/increase by 10 given that the stringency index doesn't exceed 100 or go below 0.
    \item Transition dynamics $\altmathcal{T}\left(\mathbf{s}_{t+1} \mid \mathbf{s}_t, \mathbf{a}_t\right)$ that map a state-action pair at time $t$ onto a distribution of states at time $t+1$. This state transition is defined by the SIR model with lockdown and the model of how stringency index affects the GDP.
    \item Immediate reward $\altmathcal{R}\left(\mathbf{s}_t, \mathbf{a}_t, \mathbf{s}_{t+1}\right)$. The agent observes the state of the environment $\mathbf{s}_t$ at time $t$ and takes an action $\mathbf{a}_t$, after which the state transitions to $\mathbf{s}_{t+1}$ and the agent receives a reward $\mathbf{r}_{t+1}$ as feedback. In \cref{defining_reward_function} we define a reward strategy however it should be noted that this work serves as a framework where the strategy can be easily swapped for another to prioritize different needs.
    \item Discount Factor $\gamma \in[0,1]$, where lower values place more emphasis on immediate rewards. Here, we choose the default discount factor of 0.99.
\end{itemize}

In general, the policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$. If the MDP is episodic, i.e., the state is reset after each episode of length $T$, then the sequence of states, actions and rewards in an episode constitutes a trajectory or rollout of the policy. Every rollout of a policy accumulates rewards from the environment, resulting in the return $R=\sum_{t=0}^{t=T-1} \gamma^t r_{t+1}$. The goal of RL is to find an optimal policy, $\pi^*$, which achieves the maximum expected return from all states. To achieve this, reinforcement learning start with an initial arbitrary policy, i.e., a $Q$-table with no entries. $Q$-table is a mapping from states $s_t \in \mathscr{S}$ to a predefined set of actions to increase or decrease the stringency at time t, which are the actions $a_t \in \mathscr{A}$. Each entry of the $Q$-table $\left(Q_t\left(s_t, a_t\right)\right)$ associates an action in the finite sequence $\left(\mathscr{A}_j\right)_{j \in \mathbb{J}^{+}}$ to a state of the finite sequence $\left(\mathscr{S}_i\right)_{i \in \mathbb{I}^{+}}$~\cite{sutton2018reinforcement}.

In this case of epidemic control by non-pharmaceutical interventions (NPI) based strategies this policy represents the series of stringencies to be imposed upon the population to shift the initial status of the environment to a targeted status which is equivalent to the desired set of system states. This is how the $Q$-table updates saying, if in state $s_k$ the most ideal action is $a_t$. After having more and more experience with the environment and understanding which actions lead to a higher reward $r$ an optimal policy is derived be maximizing the expected value of discounted reward $J\left(r_t\right)=\mathbb{E}\left[\sum_{t=1}^{\infty} \gamma^{(t-1)} r_t\right]$, where, discount factor $\gamma \in [0, 1]$ and time steps $k = 1, 2, \dots$.

\subsubsection{Defining the Reward Function}\label{defining_reward_function}
The stringency index emerges as a critical factor influencing both the Gross Domestic Product (GDP) and the rate of infection spread. The decision to escalate or de-escalate the stringency index is a strategic one, with significant implications. Increasing the stringency decreases the spread of the infection. Conversely, it must be noted that herd immunity can only be achieved when the epidemic reaches its peak i.e. when the effective reproductive number is equal to one ($R_e = 1$). This can only happen by lowering the stringency index which would allow the natural dynamics of the epidemic to transpire such that the population of susceptible individuals has depleted enough such that it is insufficient to propagate the disease further. Therefore, stringency is used to control the number of infected people and slow down the rate at which the epidemic reaches its peak, so that hospitals could house the number of infected people.

With this basis we define the reward function~\cite{Ohi2020}. In Deep Reinforcement Learning (DRL), positive rewards promote and negative rewards demote actions. The agent tries to generate such a policy/knowledge to avoid the discouraging situation by following the policy. By designing a proper reward function, it is possible to generate such an agent that may follow the human desired situation. While designing a reward function it is important to note that the rewards we set up truly indicate what we want accomplished. In particular, the reward signal is not the place to impart to the agent prior knowledge about how to achieve what we want it to do~\cite{sutton2018reinforcement}.
% TODO add: http://www.incompleteideas.net/book/ebook/node29.html

The reward function is parameterized to account for key factors influencing decision-making. To incentivize reduction of $R_e$ (effective reproductive number) the reward is $((R_e^{2}) \times -100) + 100$, this means when $R_e = 1$ there is no reward, but it exponentially decreases when $R_e > 1$ and increases if $R_e < 1$. Similarly, for the normalized GDP the reward function is $((\textrm{min\_max\_normalized\_GDP} ** 5)*300) - 80$ which tends to become negative below the average

If the proportion of the infected population were to rise above 0.003 the model is punished ($-5000$) and otherwise rewarded ($+20$). To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| * -20$).
